---
title: "Assignment-3"
author: "Julia Ferris"
date: "2023-09-21"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

## Introduction

In this document, the four parts of Assignment 3 will be answered with code as the answer or as examples.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 1: Identify the majors that contain DATA or STATISTICS.

In this section, the readr package is used to import data from a CSV file that is located in a Git Hub repository. Then, majors with DATA or STATISTICS are filtered out of the original data frame.

```{r part-1}
#Import data
library(readr)
library(stringr)
library(gt)
majors <- read_csv("https://raw.githubusercontent.com/juliaDataScience-22/cuny-fall-23/manage-acquire-data/majors-list.csv")

gt(head(majors)) |>   
  tab_header(     
    title = "Table 1",
    subtitle = "All Majors"
    )

#Filter out majors with DATA
dataMajors <- majors[str_detect(majors$Major, "DATA"),]

#Filter out majors with STATISTICS
statisticsMajors <- majors[str_detect(majors$Major, "STATISTICS"),]

newMajors <- rbind(dataMajors, statisticsMajors)

#Print information about all majors containing DATA or STATISTICS
gt(head(newMajors)) |>   
  tab_header(     
    title = "Table 2",
    subtitle = "Majors Containing DATA or STATISTICS"
    )

```

## Part 2: Transform the data into a new form.

Next, the list of items in foodItems needs to be transformed into the format it was used to create the list.

```{r part-2}
foodItems <- c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

noquote(paste0("c(",'"',foodItems[1],'"',", ",'"',foodItems[2],'"',', ','"',foodItems[3],'"',', ','"',foodItems[4],'"',', ','"',foodItems[5],'"',', ','"',foodItems[6],'"',', ','"',foodItems[7],'"',', ','"',foodItems[8],'"',', ','"',foodItems[9],'"',', ','"',foodItems[10],'"',', ','"',foodItems[11],'"',', ','"',foodItems[12],'"',', ','"',foodItems[13],'"',', ','"',foodItems[14],'"',')'))

```

## Part 3: Describe what the expressions will match.

(a) (.)\\1\\1

    This expression will match with all words that contain a triple
    letter. For example, "treee" will match because of the triple E,
    while "reference" will not match because not all the E's are next to
    each other.

(b) "(.)(.)\\2\\1"

    This expression will match with all words that contain the following
    pattern: abba. The first symbol must be repeated as the fourth
    symbol. The middle two symbols must match each other. For example,
    "sees" will match, but "keep" will not.

(c) (..)\\1

    This expression will match with all words that contain a repeated
    pattern of two symbols. It will match with the following pattern:
    abab. For example, "banana" will match because it contains anan,
    while "decide" will not match because de and de do not repeat next
    to each other.

(d) "(.).\\1.\\1"

    This expression will match with all words that contain the same
    symbol three times as every other symbol. The pattern is abaca where
    b and c can be anything. For example, "caravan" will match because
    of the arava part, while "greeter" will not match because the three
    E's are not every other letter.

(e) "(.)(.)(.).\*\\3\\2\\1"

    This expression will match with all words in the following pattern:
    abcdecba. The first three symbols of the pattern must be the last
    three symbols in reverse order. The middle two symbols can be
    anything. For example, "snellens" and "acegheca" will match, while
    "grilling" and "automation" will not match.

All examples are provided in the code below:

```{r pressure}
words <- c("one","book","treee","banana","hello","sees","decide","snellens","abcdecba","tacos","sweetest","cucumber")

str_view(words, "(.)\\1\\1")
str_view(words, "(.)(.)\\2\\1")
str_view(words, "(..)\\1")
str_view(words, "(.).\\1.\\1")
str_view(words, "(.)(.)(.).*\\3\\2\\1")
```

## Part 4: Match words for the three situations.

(a) Start and end with the same character

    "\^(.).\*\\\\1\$"

(b) Contain a repeated pair of letters (together, like ab ab)

    "([a-z]\|[A-Z])(a-z]\|[A-Z]).\*?\\\\1\\\\2"

(c) Contain one letter repeated in at least three places

    "([a-z]\|[A-Z]).*?\\\\1.*?\\\\1"

Examples of each are provided in the following code:

```{r part-4}
str_view(words, "^(.).*?\\1$")
str_view(words, "([a-z]|[A-Z])([a-z]|[A-Z]).*?\\1\\2")
str_view(words, "([a-z]|[A-Z]).?\\1.?\\1")
```
## Sources

1. Wickham, Hadley. Çetinkaya-Rundel, Mine. Grolemund, Garrett. R for Data Science, 2nd Edition. 2023. https://r4ds.hadley.nz/

2. Lander, Jared P. R for Everyone: Advanced Analytics and Graphics. Addison Wesley, 2017.